**OptiFit: a fast method for fitting amplicon sequences to existing OTUs**

### 2021-03-25 exploratory/exploratory.Rmd

```{r setup_environment, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

@@

models <- c("rf","glmnet","xgbTree","svmRadial","rpart2")
levels <- c("phylum","class","order","family","genus","otu","asv")
levels_names <- c("Phylum","Class","Order","Family","Genus","OTU","ASV")
pal <- park_palette("Everglades",length(models))
names(pal) <- models
pal2 <- park_palette("Everglades",length(models))
names(pal2) <- c("Random Forest","Logistic Regression","XGBoost","SVM Radial","Decision Tree")

full_df <- NULL
auc_df <- NULL
cv_table <- tibble(level=c("phylum","class","order","family","genus","otu","asv"))
t_table <- tibble(level=c("phylum","class","order","family","genus","otu","asv"))

for( m in models ){
  phylum <- read_csv(paste0("./data/final/process/combined-phylum-",m,".csv")) %>% 
    mutate(level="phylum")
  class <- read_csv(paste0("./data/final/process/combined-class-",m,".csv")) %>% 
    mutate(level="class")
  order <- read_csv(paste0("./data/final/process/combined-order-",m,".csv")) %>% 
    mutate(level="order")
  family <- read_csv(paste0("./data/final/process/combined-family-",m,".csv")) %>% 
    mutate(level="family")
  genus <- read_csv(paste0("./data/final/process/combined-genus-",m,".csv")) %>%
    mutate(level="genus")
  otu <- read_csv(paste0("./data/final/process/combined-otu-",m,".csv")) %>%
    mutate(level="otu")
  asv <- read_csv(paste0("./data/final/process/combined-asv-",m,".csv")) %>%
    mutate(level="asv")

@@

#compare two models levels within a taxonomic level
p.table.level <- read_csv("./data/analysis/exploratory3/pvalues_by_level.csv")
#compare two taxonomic levels within a model type
p.table.model <- read_csv("./data/analysis/exploratory3/pvalues_by_model.csv")

@@

>>>>

```{r setup_environment, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message=FALSE)

@@

# Variables
models <- c("rf","glmnet","xgbTree","svmRadial","rpart2")
levels <- c("phylum","class","order","family","genus","otu","asv")
levels_names <- c("Phylum","Class","Order","Family","Genus","OTU","ASV")
pal <- park_palette("Everglades",length(models))
names(pal) <- models
pal2 <- park_palette("Everglades",length(models))
names(pal2) <- c("Random Forest","Logistic Regression","XGBoost","SVM Radial","Decision Tree")

# Generate dataframe of all results and tables of the median cv and test AUCs
full_df <- NULL
auc_df <- NULL
cv_table <- tibble(level=c("phylum","class","order","family","genus","otu","asv"))
t_table <- tibble(level=c("phylum","class","order","family","genus","otu","asv"))

for( m in models ){
  phylum <- read_csv(paste0("../data/process/combined-phylum-",m,".csv")) %>% 
    mutate(level="phylum")
  class <- read_csv(paste0("../data/process/combined-class-",m,".csv")) %>% 
    mutate(level="class")
  order <- read_csv(paste0("../data/process/combined-order-",m,".csv")) %>% 
    mutate(level="order")
  family <- read_csv(paste0("../data/process/combined-family-",m,".csv")) %>% 
    mutate(level="family")
  genus <- read_csv(paste0("../data/process/combined-genus-",m,".csv")) %>%
    mutate(level="genus")
  otu <- read_csv(paste0("../data/process/combined-otu-",m,".csv")) %>%
    mutate(level="otu")
  asv <- read_csv(paste0("../data/process/combined-asv-",m,".csv")) %>%
    mutate(level="asv")

@@

write_csv(t_table,"./figures/median_test_auc.csv")

#compare two models levels within a taxonomic level
p.table.level <- read_csv("../analysis/pvalues_by_level.csv")
#compare two taxonomic levels within a model type
p.table.model <- read_csv("../analysis/pvalues_by_model.csv")

@@

## Input data overview

To total number of samples and features before and after preprocessing for each taxnomic level. 
```{r input}

input <- read_csv("../analysis/input_values.csv")

kable(input)
```

## Hyperparameter Performance

Plots of HP performance for each model/taxonomic level

### Random Forest
The default hyperparameter (mtry) selection for RF is:

1. sqrt_features / 2
1. sqrt_features
1. sqrt_features * 2

Based on previous results, I added mtry=1 for Phylum through Genus levels and mtry=100 for OTU and ASV levels. This set of hyperparameters seems to cover the optimal range for each taxonomic level.

![](../analysis/hp-rf.png)

### Logistic Regression

Two metrics: alpha and lambda  
By default alpha is set to zero (for L2 regularization) and lambda values are: 1e-04, 1e-03, 1e-02, 1e-01, 1e+00, 1e+01
I added 1e-05 as a lambda value for phylum and class, and higher values of lambda for the rest

![](../analysis/hp-glmnet.png)

### Decision Tree

The default hyperparameters were retained for decision tree (maxdepth = 1  2  4  8 16 30). It would not allow larger maxdepth than 30.

![](../analysis/hp-rpart2.png)

### XGBoost

(needs adjustment)

### SVM Radial

![](../analysis/hp-svmRadial.png)

## Model Performance

```{r plot_by_method, echo=F,message=F,fig.height = 3.5, fig.width = 10, fig.align = "center"}

df_by_method <- full_df %>% 
  select(cv_metric_AUC,AUC,method,level) %>% 
  rename(test_AUC=AUC,cv_AUC=cv_metric_AUC) %>% 
  pivot_longer(cols=c("cv_AUC","test_AUC"),names_to="auc_type",values_to="AUC")

df_by_method$method <- factor(df_by_method$method,
                        levels=c("rf","glmnet","xgbTree","svmRadial","rpart2"),
                        labels=c("Random Forest","Logistic Regression","XGBoost","SVM Radial","Decision Tree"))

ggplot(df_by_method,aes(x=factor(level,
                            levels=c("phylum","class","order","family","genus","otu","asv"),
                            labels=c("Phylum","Class","Order","Family","Genus","OTU","ASV")),
                   y=AUC, fill=auc_type)) + 
    geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
    geom_boxplot(notch = TRUE) + 
    facet_grid(.~method) + 
    theme_bw() + xlab("") + ylab("AUROC") +
    theme(legend.position ="top",
          axis.text.x = element_text(angle = 45,hjust = 1)) 
ggsave("./figures/performance_by_method.png")
```

```{r plot_by_level, echo=FALSE, message=FALSE,fig.height = 6, fig.width = 10, fig.align = "center"}
df_by_level <- full_df %>% 
  select(cv_metric_AUC,AUC,method,level) %>% 
  rename(test_AUC=AUC,cv_AUC=cv_metric_AUC) %>% 
  pivot_longer(cols=c("cv_AUC","test_AUC"),names_to="auc_type",values_to="AUC") %>% 
  mutate(level=factor(level,levels=c("phylum","class","order","family","genus","otu","asv"),
                            labels=c("Phylum","Class","Order","Family","Genus","OTU","ASV")),
         method=factor(method,levels=c("rf","glmnet","xgbTree","svmRadial","rpart2"),
                       labels=c("Random Forest","Logistic Regression","XGBoost","SVM Radial","Decision Tree")))

ggplot(df_by_level,aes(x=method, y=AUC, fill=auc_type)) + 
  geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
  geom_boxplot(notch = TRUE) + 
  facet_wrap(.~level,nrow = 2) + 
  theme_bw() + xlab("") + ylab("AUROC") + 
  theme(legend.position ="top",
        axis.text.x = element_text(angle = 45,hjust = 1)) 
ggsave("./figures/performance_by_tax.png")

```{r plot_all, echo=F, message=F, out.width="100%",fig.height=4}
df_all <- full_df %>% 
  select(cv_metric_AUC,AUC,method,level) %>% 
  rename(test_AUC=AUC,cv_AUC=cv_metric_AUC) %>% 
  pivot_longer(cols=c("cv_AUC","test_AUC"),names_to="auc_type",values_to="AUC") %>% 
  mutate(level=factor(level,levels=c("phylum","class","order","family","genus","otu","asv"),
                            labels=c("Phylum","Class","Order","Family","Genus","OTU","ASV")),
         method=factor(method,levels=c("rf","glmnet","xgbTree","svmRadial","rpart2"),
                       labels=c("Random Forest","Logistic Regression","XGBoost","SVM Radial","Decision Tree")))

df_all %>% 
  filter(auc_type == "test_AUC") %>% 
  ggplot(aes(x=level,y=AUC,fill=method)) + 
    geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
    geom_boxplot(notch = TRUE) + 
    theme_bw() + xlab("") + ylab("AUROC") + 
    theme(legend.position ="top",
          axis.text.x = element_text(angle = 45,hjust = 1)) +
    scale_fill_manual(values=pal2,name="Model")
ggsave("./figures/all_model_level.png")
```

### DADA2 Comparison

```{r dada2}

dada2_results <- read

### Random Forest Model Performance with Significance

Random Forest on OTU level data yields the highest median AUC, followed by Family, Genus, and ASV. While the OTU AUC is not significantly higher than that of Family or Genus level, it is significantly higher than ASV.

<div class = "row">
<div class = "col-md-6">
```{r rf_table,echo=FALSE}
t_table %>%
  select(level,rf) %>% 
  mutate(rf=round(rf,digits = 3)) %>% 
  rename("Median AUC"=rf,"Level"=level) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position="center")
```

</div>

<div class = "col-md-6">
```{r rf_plot,echo=F,message=F}
m <- "Random Forest"

sub_df <- auc_df %>% 
  mutate(level=factor(level,levels=c("phylum","class","order","family","genus","otu","asv"),
                       labels=c("Phylum","Class","Order","Family","Genus","OTU","ASV")),
         method=factor(method,levels=c("rf","glmnet","xgbTree","svmRadial","rpart2"),
                       labels=c("Random Forest","Logistic Regression","XGBoost","SVM Radial","Decision Tree"))) %>% 
  filter(method==m & auc_type=="test_AUC")

sub_stats <- p.table.model %>% 
   mutate(level1=factor(level1,levels=c("phylum","class","order","family","genus","otu","asv"),
                            labels=c("Phylum","Class","Order","Family","Genus","OTU","ASV")),
          level2=factor(level2,levels=c("phylum","class","order","family","genus","otu","asv"),
                            labels=c("Phylum","Class","Order","Family","Genus","OTU","ASV")),
         model=factor(model,levels=c("rf","glmnet","xgbTree","svmRadial","rpart2"),
                       labels=c("Random Forest","Logistic Regression","XGBoost","SVM Radial","Decision Tree"))) %>% 
  filter(model==m) %>% 
  mutate(sig=case_when(p_value >= 0.05 ~ "NS",
                       TRUE ~ "*")) %>% 
  filter(sig=="NS")

sub_stats <- sub_stats %>% 
  mutate(y.position=seq.int(0.85,1.05,length.out = nrow(sub_stats))) %>% 
  rename(group1=level1,group2=level2)

ggplot(sub_df,aes(x=level,y=AUC)) + 
  geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
  geom_boxplot(aes(fill=method),notch = T) +
  scale_fill_manual(values = pal2[m]) + 
  scale_y_continuous(limits=c(0.2,1.05)) +
  theme_bw() + xlab("") + ylab("AUROC") +
  theme(legend.position ="none",
        axis.text.x = element_text(angle = 45,hjust = 1),
        panel.grid.major.x = element_blank(),
        axis.text = element_text(size=14)) +
  ggtitle(m) +
  stat_summary(fun = mean, color="white",geom = "point", shape = 18, size = 3,show.legend = FALSE) +
  stat_pvalue_manual(sub_stats,label="sig")
ggsave("./figures/rf_with_stats.png",width = 5,height=4,units="in")

```
</div>
</div>

### Logistic Regression Model Performance with Significance

<div class = "row">
<div class = "col-md-6">
```{r glmnet_table,echo=FALSE}
t_table %>%
  select(level,glmnet) %>% 
  mutate(glmnet=round(glmnet,digits = 3)) %>% 
  rename("Median AUC"=glmnet,"Level"=level) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position="center")
```
</div>

<div class = "col-md-6">
```{r glmnet_plot,echo=F,message=F}
## glmnet Model Performance
m <- "Logistic Regression"

sub_df <- auc_df %>% 
  mutate(level=factor(level,levels=c("phylum","class","order","family","genus","otu","asv"),
                       labels=c("Phylum","Class","Order","Family","Genus","OTU","ASV")),
         method=factor(method,levels=c("rf","glmnet","xgbTree","svmRadial","rpart2"),
                       labels=c("Random Forest","Logistic Regression","XGBoost","SVM Radial","Decision Tree"))) %>% 
  filter(method==m & auc_type=="test_AUC")

sub_stats <- p.table.model %>% 
   mutate(level1=factor(level1,levels=c("phylum","class","order","family","genus","otu","asv"),
                            labels=c("Phylum","Class","Order","Family","Genus","OTU","ASV")),
          level2=factor(level2,levels=c("phylum","class","order","family","genus","otu","asv"),
                            labels=c("Phylum","Class","Order","Family","Genus","OTU","ASV")),
         model=factor(model,levels=c("rf","glmnet","xgbTree","svmRadial","rpart2"),
                       labels=c("Random Forest","Logistic Regression","XGBoost","SVM Radial","Decision Tree"))) %>% 
  filter(model==m) %>% 
  mutate(sig=case_when(p_value >= 0.05 ~ "NS",
                       TRUE ~ "*")) %>% 
  filter(sig=="NS")

sub_stats <- sub_stats %>% 
  mutate(y.position=seq.int(0.85,1.05,length.out = nrow(sub_stats))) %>% 
  rename(group1=level1,group2=level2)

ggplot(sub_df,aes(x=level,y=AUC)) + 
  geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
  geom_boxplot(aes(fill=method),notch = T) +
  scale_fill_manual(values = pal2[m]) + 
  scale_y_continuous(limits=c(0.2,1.05)) +
  theme_bw() + xlab("") + ylab("AUROC") +
  theme(legend.position ="none",
        axis.text.x = element_text(angle = 45,hjust = 1),
        panel.grid.major.x = element_blank(),
        axis.text = element_text(size=14)) +
  ggtitle(m) +
  stat_summary(fun = mean, color="white",geom = "point", shape = 18, size = 3,show.legend = FALSE) +
  stat_pvalue_manual(sub_stats,label="sig")
ggsave("./figures/glmnet_with_stats.png",width = 5,height=4,units="in")

```
</div>
</div>

### Decision Tree Model Performance with Significance

<div class = "row">
<div class = "col-md-6">
```{r rpart2_table,echo=FALSE}
t_table %>%
  select(level,rpart2) %>% 
  mutate(rpart2=round(rpart2,digits = 3)) %>% 
  rename("Median AUC"=rpart2,"Level"=level) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position="center")
```
</div>

<div class = "col-md-6">
```{r rpart2_plot,echo=F,message=F}
## rpart2 Model Performance
m <- "Decision Tree"

sub_df <- auc_df %>% 
  mutate(level=factor(level,levels=c("phylum","class","order","family","genus","otu","asv"),
                       labels=c("Phylum","Class","Order","Family","Genus","OTU","ASV")),
         method=factor(method,levels=c("rf","glmnet","xgbTree","svmRadial","rpart2"),
                       labels=c("Random Forest","Logistic Regression","XGBoost","SVM Radial","Decision Tree"))) %>% 
  filter(method==m & auc_type=="test_AUC")

sub_stats <- p.table.model %>% 
   mutate(level1=factor(level1,levels=c("phylum","class","order","family","genus","otu","asv"),
                            labels=c("Phylum","Class","Order","Family","Genus","OTU","ASV")),
          level2=factor(level2,levels=c("phylum","class","order","family","genus","otu","asv"),
                            labels=c("Phylum","Class","Order","Family","Genus","OTU","ASV")),
         model=factor(model,levels=c("rf","glmnet","xgbTree","svmRadial","rpart2"),
                       labels=c("Random Forest","Logistic Regression","XGBoost","SVM Radial","Decision Tree"))) %>% 
  filter(model==m) %>% 
  mutate(sig=case_when(p_value >= 0.05 ~ "NS",
                       TRUE ~ "*")) %>% 
  filter(sig=="NS")

sub_stats <- sub_stats %>% 
  mutate(y.position=seq.int(0.85,1.05,length.out = nrow(sub_stats))) %>% 
  rename(group1=level1,group2=level2)

ggplot(sub_df,aes(x=level,y=AUC)) + 
  geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
  geom_boxplot(aes(fill=method),notch = T) +
  scale_fill_manual(values = pal2[m]) + 
  scale_y_continuous(limits=c(0.2,1.05)) +
  theme_bw() + xlab("") + ylab("AUROC") +
  theme(legend.position ="none",
        axis.text.x = element_text(angle = 45,hjust = 1),
        panel.grid.major.x = element_blank(),
        axis.text = element_text(size=14)) +
  ggtitle(m) +
  stat_summary(fun = mean, color="white",geom = "point", shape = 18, size = 3,show.legend = FALSE) +
  stat_pvalue_manual(sub_stats,label="sig")
ggsave("./figures/rpart2_with_stats.png",width = 5,height=4,units="in")

```
</div>
</div>

### SVM Radial

### XGBTree

## Feature Importance
The mikropml packages includes an option for finding [feature importance](http://www.schlosslab.org/mikropml/reference/get_feature_importance.html) using a permutation method. This significantly increases the time to run the models. The function permutes each model feature (e.g. genus) and recalulates the AUC with that feature permuted. The outputs are the permuted AUC value and the difference between this AUC value and the actual AUC value. Larger positive differences can be interpreted as being more important since the AUC value is smaller when this feature is permuted.

The importance values are all from the Random Forest model.

### 2021-03-25 paper/manuscript.Rmd

Courtney R. Armour, (ande? begum?), Patrick D. Schloss ${^\dagger}$

@@

- Across the five ML methods tested, model performance tended to increase with taxonomic level usually peaking around genus/OTU level before dropping off slightly with ASVs. [(Figure)](#model-performance-across-taxonomy)
- Random forest was consistently the top performer at most taxonomic levels.
  - RF might be more appropriate for microbiome analysis since its more suitable for zero inflated data 
  - TODO: dig into literature on this
  - TODO: address lower L2 logistic performance that prior results?
- Within the RF model, the highest AUCs were observed for family (median AUC: `r round(rf_med_auc["family"],digits=3)`), genus(median AUC: `r round(rf_med_auc["genus"],digits=3)`), and OTU (median AUC: `r round(rf_med_auc["otu"],digits=3)`) level data with no significant difference between the three. [(Figure)](#random-forest-model-performance-with-significance) 
  - ASV (median AUC: `r round(rf_med_auc["asv"],digits=3)`) performed significantly lower than OTU and genus (p < 0.05) but not family level P = 0.06).
- One hypothesis for the observation that model performance increases from phylum to OTU level then drops slightly at ASV level is that at higher taxonomic levels (e.g. phylum) there are too few taxa and too much overlap to reliably differentiate between cases and controls. 
  - As you reach genus/OTU level data there is enough data and variation but at the ASV level, the data is too specific to individuals and doesn't overlap enough. 
  - Examination of the prevalence of taxa in samples at each level supports this idea. A majority of taxa are present in greater than 75% of samples at the phylum (67% of taxa) and class (63% of taxa) levels.  The opposite is observed at the OTU and ASV level where 60% and 53% of taxa respectively are only present in less than 25% of samples. [(Figure)](#prevalence-of-taxa-in-samples)
- Of note, the ML pipeline includes a pre-processing step that occurs prior to training and classifying the ML models (methods). As part of this step, features are removed that wont provide useful information to build the model. For example, strongly correlated features provide the same information to build the model and thus can be collapsed. Additionally, features with zero or near-zero variance will not help the model differentiate groups and thus can be removed.
  - Interestingly, despite starting with `r input %>% filter(level == "asv") %>% select(n_features) %>% pull()` features at the ASV level, only `r input %>% filter(level == "asv") %>% select(n_features_preproc) %>% pull()` (`r input %>% filter(level == "asv") %>% select(pct_kept) %>% pull()`%) remained after pre-processing. At the OTU level, `r input %>% filter(level == "otu") %>% select(n_features) %>% pull()` of the `r input %>% filter(level == "otu") %>% select(n_features_preproc) %>% pull()` features (`r input %>% filter(level == "otu") %>% select(pct_kept) %>% pull()`%) remained after preprocessing. [(Table)](#summary-of-features)
  - TODO: check why ASVs were removed in preprocessing (nzv?)
  - (depending on above can discuss why features were removed - e.g. only in one or a few samples)
- OTU level data provides an idea balance of overlap and distinction to classify samples. 
- Overall, the fine resolution of ASVs is unnecessary for ML classification with microbiome data since the vast majority of the ASVs are removed during preprocessing and the model performance is diminshed compared to other taxonomic levels.

@@

## Materials and Methods
- 16S rRNA data from 490 subjects {baxter}
  - 261 controls
  - 229 cases
- processed with mothur v1.44.3
  - SILVA v132 reference
- ML with mikropml package
  - preprocessing
    -normalize values
    - what is removed and why
      - near zero variance & zero variance removed
      - correlated collapsed
  - hp tuning
  - n seeds
  - for more detail see mikropml package{}
- pvalues calculated as previously described {begum}
- prevalence = number of samples with non-zero abundance / total number of samples

>>>>

Courtney R. Armour, Begüm D.Topçuoğlu, Andrea Garretto, Patrick D. Schloss ${^\dagger}$

@@

- Across the five ML methods tested, model performance tended to increase with taxonomic level usually peaking around genus/OTU level before dropping off slightly with ASVs. [(Figure)](#model-performance-across-taxonomy)
- Random forest was consistently the top performer at most taxonomic levels.
  - RF might be more appropriate for microbiome analysis since its more suitable for zero inflated data 
  - TODO: dig into literature on this
- Within the RF model, the highest AUCs were observed for family (median AUC: `r round(rf_med_auc["family"],digits=3)`), genus(median AUC: `r round(rf_med_auc["genus"],digits=3)`), and OTU (median AUC: `r round(rf_med_auc["otu"],digits=3)`) level data with no significant difference between the three. [(Figure)](#random-forest-model-performance-with-significance) 
  - ASV (median AUC: `r round(rf_med_auc["asv"],digits=3)`) performed significantly lower than OTU and genus (p < 0.05) but not family level P = 0.06).
  - TODO: compare dada2 performance
- One hypothesis for the observation that model performance increases from phylum to OTU level then drops slightly at ASV level is that at higher taxonomic levels (e.g. phylum) there are too few taxa and too much overlap to reliably differentiate between cases and controls. 
  - As you reach genus/OTU level data there is enough data and variation but at the ASV level, the data is too specific to individuals and doesn't overlap enough. 
  - Examination of the prevalence of taxa in samples at each level supports this idea. A majority of taxa are present in greater than 75% of samples at the phylum (67% of taxa) and class (63% of taxa) levels.  The opposite is observed at the OTU and ASV level where 60% and 53% of taxa respectively are only present in less than 25% of samples. [(Figure)](#prevalence-of-taxa-in-samples)
- Of note, the ML pipeline includes a pre-processing step that occurs prior to training and classifying the ML models (methods). As part of this step, features are removed that wont provide useful information to build the model. For example, strongly correlated features provide the same information to build the model and thus can be collapsed. Additionally, features with zero or near-zero variance will not help the model differentiate groups and thus can be removed.
  - Interestingly, despite starting with `r input %>% filter(level == "asv") %>% select(n_features) %>% pull()` features at the ASV level, only `r input %>% filter(level == "asv") %>% select(n_features_preproc) %>% pull()` (`r input %>% filter(level == "asv") %>% select(pct_kept) %>% pull()`%) remained after pre-processing. At the OTU level, `r input %>% filter(level == "otu") %>% select(n_features) %>% pull()` of the `r input %>% filter(level == "otu") %>% select(n_features_preproc) %>% pull()` features (`r input %>% filter(level == "otu") %>% select(pct_kept) %>% pull()`%) remained after preprocessing. [(Table)](#summary-of-features)
  - TODO: check why ASVs were removed in preprocessing (nzv?)
  - (depending on above can discuss why features were removed - e.g. only in one or a few samples)
- OTU level data provides an idea balance of overlap and distinction to classify samples. 
- Overall, the fine resolution of ASVs is unnecessary for ML classification with microbiome data since the vast majority of the ASVs are removed during preprocessing and the model performance is diminshed compared to other taxonomic levels.

@@

## Materials and Methods
- 16S rRNA data from 490 subjects {baxter}
  - 261 controls
  - 229 SRNs (describe how SRN defined)
- processed with mothur v1.44.3
  - SILVA v132 reference
- ML with mikropml package
  - preprocessing
    -normalize values
    - what is removed and why
      - near zero variance & zero variance removed
      - correlated collapsed
  - hp tuning
  - n seeds
  - for more detail see mikropml package{}
- pvalues calculated as previously described {begum}
- prevalence = number of samples with non-zero abundance / total number of samples
